---
title: "Diagnostics and Uncertainty Boundaries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Diagnostics and Uncertainty Boundaries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## Inspect every fit

```{r diagnostics}
print(fit)
fit$diagnostics
```

The diagnostics contract contains:

- `n_divergent`;
- `max_treedepth_configured`, `max_treedepth_observed`, and `max_depth_hits`;
- `rhat_max`, `rhat_all_ok`, and `rhat_n` for structural parameters;
- `n_eff_min`, `n_eff_median`, and `n_eff_n`.

The last fields are explicitly labeled legacy rstan `n_eff`; they are not
bulk or tail ESS. `bhf_fit()` warns at any divergence, R-hat at least 1.05,
legacy `n_eff` below 100, or a configured tree-depth hit. Those warnings are
screening rules, not proof that an otherwise quiet fit is scientifically valid.

Tune through the named public arguments:

```{r tuning}
fit <- bhf_fit(
  prepared,
  model = model,
  chains = 4,
  iter = 4000,
  warmup = 2000,
  adapt_delta = 0.99,
  max_treedepth = 15,
  seed = 1234
)
```

Do not pass a `control` list through `...`; the public function constructs it
from `adapt_delta` and `max_treedepth`.

## Pseudo-posterior intervals

The fitted target is a survey-weighted pseudo-likelihood. R-hat and effective
sample size concern numerical exploration of that target; they do not establish
frequentist coverage or correct design-based uncertainty. Sandwich adjustment
is not implemented. Interpret credible intervals with this limitation.

## Fixed-vhat limitation

Taylor or supplied domain sampling variances enter A* as fixed values. Their
estimation uncertainty is not propagated. Always retain
`fit$provenance$sampling_variances` and `vd$A_star$provenance` with results.

## Log-likelihood scope

```{r loglik}
ll <- log_lik(fit, kind = "pseudo", aggregate = "observation")
attr(ll, "scope")
```

The pseudo form emits a warning. `aggregate = "psu"` or `"stratum"` sums
stored contributions; it does not refit after leaving out a unit and does not
establish ordinary cluster-level LOO or WAIC validity. The raw form is an
unweighted Bernoulli diagnostic, not the fitted target.

## What diagnostics do and do not tell you

Sampler diagnostics answer one question: did the sampler explore the posterior
properly? They do not tell you whether the resulting intervals have the
frequentist coverage you might expect.

In the 400-fit reduced balanced synthetic study, sampler health and repeated-
sample coverage were evaluated separately. Pooled primary 90% coverage was
0.900, while the four fixed-condition rates ranged from 0.860 to 0.930. Sampler
health is necessary but not sufficient: treat clean diagnostics as a
precondition for interpreting a fit, not as evidence about calibration.
