---
title: "Introduction to bhfvar"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to bhfvar}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## Purpose

`bhfvar` is an experimental implementation of a Bayesian pseudo-likelihood
framework for decomposing binary-outcome variation across post-hoc domains in a
complex survey. It compares a domain-only reference standardization with a
standardization that retains fitted stratum and PSU-within-stratum effects.

The package addresses two distinct questions:

1. What probability-scale variation is induced by domain effects alone (A)?
2. What variation is present after retaining the observed mixture of design
   effects within each domain (B)?

A* subtracts a fixed, population-share-weighted sampling-variance correction
from A's between-domain component, truncated at zero.

## Model structure

For observation $i$,

$$
\operatorname{logit}(p_i)=\alpha+u_{s[i]}+v_{h[i]}+b_{j[i]},
$$

where domain and stratum effects are crossed and each flat PSU identifier is
nested in exactly one stratum. State effects are centered using population
shares, stratum effects by their simple mean, and PSU effects within stratum.

```text
observations
 ├─ domain/state (crossed)
 └─ stratum
     └─ PSU (nested)
```

## Supported workflow

```{r workflow}
library(bhfvar)
data(bhf_synthetic_data)

prepared <- prepare_bhf_data(
  bhf_synthetic_data,
  outcome = "has_subsidy",
  domain = "state",
  strata = "stratum",
  psu = "psu",
  weights = "weight",
  deattenuation = "taylor"
)
model <- compile_bhf_model()
fit <- bhf_fit(prepared, model = model, chains = 4, iter = 2000,
               seed = 1234)

variance_decomposition(fit)
domain_estimates(fit, estimand = "A")
domain_estimates(fit, estimand = "B")
```

The example is intentionally not evaluated while building the vignette because
Stan compilation and MCMC require a configured toolchain. The data-only path is
evaluated in the quick-start vignette and dedicated tests cover public result
contracts.

## Scope

The implementation is verified against the article's main-text equations and
Appendix B Stan program: a frozen reference oracle, algebraic property tests,
and centering-invariant tests all pass.

Interval calibration is a separate question. In a reduced article-aligned
balanced synthetic design, pooled primary 90% coverage was 0.900 (95% cluster
interval 0.871 to 0.928), and fixed-condition rates ranged from 0.860 to 0.930.
That pooled result met a preregistered +/-5 percentage-point tolerance; it does
not guarantee coverage in every condition or other designs. A variance
adjustment is not implemented. The article's restricted-data application is not
reproduced by this package.

A successful compile, fit, or extraction shows that the workflow ran. It says
nothing about interval calibration. See
`vignette("scientific-limitations")`.

## Scope boundaries

- A* conditions on supplied or Taylor-estimated sampling variances; uncertainty
  in those estimates is not propagated.
- A is a design-effect-zero reference standardization, not a causal or policy
  intervention; A-versus-B is not specific to informative sampling.
- Pseudo-posterior intervals require design-aware interpretation.
- `log_lik()` does not establish ordinary observation- or cluster-level LOO.
- Plotting helpers (`bhf_plot_*()`) need `ggplot2`, a suggested dependency.
- No posterior-predictive helper is supported.
- The bundled legacy synthetic data are illustrative, not restricted NSECE
  data and not the confirmatory recovery fixture.

See `vignette("methodology", package = "bhfvar")`,
`vignette("dual-estimands", package = "bhfvar")`, and
`vignette("diagnostics", package = "bhfvar")` before substantive use.
