---
title: "Quick Start"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## A checked data-only path

This chunk is evaluated during vignette build. De-attenuation is disabled here
so the example remains lightweight and does not turn a legacy illustrative data
object into evidence about sampling-variance recovery.

```{r prepare}
library(bhfvar)
data(bhf_synthetic_data)

prepared <- prepare_bhf_data(
  bhf_synthetic_data,
  outcome = "has_subsidy",
  domain = "state",
  strata = "stratum",
  psu = "psu",
  weights = "weight",
  weight_scaling = "mean_one",
  deattenuation = "none"
)

list(
  schema = prepared$schema_version,
  observations = prepared$stan_data$N,
  domains = prepared$stan_data$S,
  strata = prepared$stan_data$H,
  nested_psus = prepared$stan_data$J
)
sum(prepared$stan_data$w_lik)
```

The final line equals the retained sample size: default likelihood weights are
globally normalized to mean one.

## Compile and fit locally

The following is not evaluated during documentation builds.

```{r fit, eval=FALSE}
model <- compile_bhf_model()
fit <- bhf_fit(
  prepared,
  model = model,
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 1234,
  adapt_delta = 0.95,
  max_treedepth = 12
)
print(fit)
fit$diagnostics
```

`fit$diagnostics` reports divergences, configured/observed tree depth, maximum
R-hat, and legacy rstan `n_eff`. It does not label `n_eff` as bulk or tail ESS.

## Extract explicit quantities

```{r extract, eval=FALSE}
vd <- variance_decomposition(fit, prob = 0.95)
vd$A$summary
vd$A_star$available
vd$A_star$provenance
vd$B$summary
vd$gaps$table

domain_estimates(fit, estimand = "A", prob = 0.95)
domain_estimates(fit, estimand = "B", prob = 0.95)
overall_estimate(fit, estimand = "B", prob = 0.95)
```

If `deattenuation = "none"`, A* remains in the result schema but
`vd$A_star$available` is `FALSE`. Re-prepare with `deattenuation = "taylor"` or
`"supplied"` to compute A* under the documented fixed-input limitation.

## Interpretation checkpoint

A workflow that runs cleanly tells you the pipeline worked; it tells you
nothing about interval calibration. The intervals returned here are
pseudo-posterior credible intervals whose frequentist coverage has not been
established. See `vignette("scientific-limitations")` before reporting them.
