---
title: "Quick Start: Your First Variance Decomposition in 5 Minutes"
author: "JoonHo Lee"
date: "`r Sys.Date()`"
description: >
  Learn how to perform your first Bayesian variance decomposition with bhfvar 
  in just 5 minutes using the included synthetic data.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start: Your First Variance Decomposition in 5 Minutes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  fig.align = "center",
  out.width = "85%",
  eval = FALSE
)
set.seed(42)
```

## Overview

This vignette demonstrates the fastest path to performing variance 
decomposition with the bhfvar package. By the end of this 5-minute tutorial, 
you will be able to:

1. Prepare complex survey data for analysis
2. Fit the Bayesian Hybrid Framework model
3. Extract and interpret variance decomposition results
4. Obtain domain-specific estimates with appropriate shrinkage


## 1. Minimal Example

Let's work through a complete analysis using the package's synthetic dataset. 
This dataset mimics the structure of the National Survey of Early Care and 
Education (NSECE), with 50 states, 27 strata, and 356 PSUs.

```{r minimal-example, eval=FALSE}
library(bhfvar)
library(rstan)

# Enable parallel processing
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)

# Step 1: Compile the Stan model (once per session)
model <- compile_bhf_model()

# Step 2: Load and prepare the data
data(bhf_synthetic_data)

prepared <- prepare_bhf_data(
  data = bhf_synthetic_data,
  outcome = "has_subsidy",
  domain = "state",
  strata = "stratum",
  psu = "psu",
  weights = "weight"
)

# Step 3: Fit the model
fit <- bhf_fit(prepared, model = model, chains = 4, iter = 2000)

# Step 4: Get results
variance_decomposition(fit)
```

That's it! You now have a complete Bayesian variance decomposition.


## 2. Understanding the Data

Let's examine the synthetic dataset:

```{r data-structure, eval=FALSE}
data(bhf_synthetic_data)

# Structure
str(bhf_synthetic_data)

# First few rows
head(bhf_synthetic_data)

# Summary statistics
cat("Total observations:", nrow(bhf_synthetic_data), "\n")
cat("Number of states:", length(unique(bhf_synthetic_data$state)), "\n")
cat("Number of strata:", length(unique(bhf_synthetic_data$stratum)), "\n")
cat("Number of PSUs:", length(unique(bhf_synthetic_data$psu)), "\n")
cat("Overall proportion with subsidy:", 
    round(mean(bhf_synthetic_data$has_subsidy), 3), "\n")
```

The dataset contains:

| Variable | Description |
|----------|-------------|
| `state` | Domain identifier (State_01 to State_50) |
| `stratum` | Design stratum (Stratum_01 to Stratum_27) |
| `psu` | Primary Sampling Unit within stratum |
| `weight` | Survey weight |
| `has_subsidy` | Binary outcome (1 = receives subsidy, 0 = does not) |


## 3. Preparing Data for Stan

The `prepare_bhf_data()` function transforms your data into the format 
required by the Stan model:

```{r data-preparation, eval=FALSE}
prepared <- prepare_bhf_data(
  data = bhf_synthetic_data,
  outcome = "has_subsidy",
  domain = "state",
  strata = "stratum",
  psu = "psu",
  weights = "weight",
  use_deattenuation = TRUE  # Enable de-attenuation (default)
)

# Print summary
print(prepared)
```

The output shows:

- Sample size and number of domains/strata/PSUs
- Outcome proportion
- Whether de-attenuation is enabled


## 4. Fitting the Model

The `bhf_fit()` function fits the Bayesian Hybrid Framework model via Stan:

```{r model-fitting, eval=FALSE}
fit <- bhf_fit(
  data = prepared,
  model = model,
  chains = 4,       # Number of MCMC chains
  iter = 2000,      # Total iterations per chain
  warmup = 1000,    # Warmup iterations (discarded)
  seed = 42         # For reproducibility
)

# Print fit summary
print(fit)
```

The printed output includes:

- Data dimensions
- Sampling settings
- Convergence diagnostics (divergences, Rhat, ESS)
- Key parameter estimates


## 5. Extracting Results

### 5.1 Variance Decomposition

The main result is the variance decomposition across three estimands:

```{r variance-decomposition, eval=FALSE}
vd <- variance_decomposition(fit)

# The function prints a formatted summary and returns detailed results
# Access specific components:
cat("ICC on logit scale (Estimand A):", 
    round(vd$logit$icc_mean, 4), "\n")
cat("ICC on probability scale (Estimand B):", 
    round(vd$prob$icc_mean, 4), "\n")
cat("ICC de-attenuated (Estimand A*):", 
    round(vd$deatten$icc_mean, 4), "\n")
```

### 5.2 Domain Estimates

Get estimates for each domain (state):

```{r domain-estimates, eval=FALSE}
# Marginal estimates (integrating out within-domain variation)
estimates <- domain_estimates(fit, type = "marginal")

# View the results
head(estimates)

# Key columns:
# - domain: Domain name
# - mean: Posterior mean probability
# - q025, q975: 95% credible interval
# - reliability: Shrinkage factor (higher = less shrinkage)
# - pop_share: Population share of this domain
```

### 5.3 Overall Estimate

Get the population-weighted overall proportion:

```{r overall-estimate, eval=FALSE}
overall <- overall_estimate(fit)

cat("Overall proportion:", round(overall$mean, 3), 
    "(95% CI:", round(overall$q025, 3), "-", round(overall$q975, 3), ")\n")
```


## 6. Quick Diagnostic Check

Always verify convergence before interpreting results:

```{r diagnostics, eval=FALSE}
# Check the fit object's diagnostics
cat("Divergent transitions:", fit$diagnostics$n_divergent, "\n")
cat("Max Rhat:", round(fit$diagnostics$rhat_max, 3), "\n")
cat("Min ESS:", fit$diagnostics$ess_min, "\n")

# Rules of thumb:
# - Divergent transitions should be 0
# - Rhat should be < 1.01 (ideally < 1.001)
# - ESS should be > 400 for reliable inference
```


## 7. Complete Workflow Summary

Here's the complete workflow in a single code block:

```{r complete-workflow, eval=FALSE}
# ============================================================
# bhfvar Quick Start: Complete Workflow
# ============================================================

library(bhfvar)
library(rstan)

# Setup
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)

# 1. Compile model
model <- compile_bhf_model()

# 2. Load data
data(bhf_synthetic_data)

# 3. Prepare data
prepared <- prepare_bhf_data(
  data = bhf_synthetic_data,
  outcome = "has_subsidy",
  domain = "state",
  strata = "stratum",
  psu = "psu",
  weights = "weight"
)

# 4. Fit model
fit <- bhf_fit(prepared, model = model, chains = 4, iter = 2000)

# 5. Results
variance_decomposition(fit)          # Variance decomposition
estimates <- domain_estimates(fit)   # Domain-specific estimates
overall <- overall_estimate(fit)     # Overall estimate

# 6. Check diagnostics
print(fit)
```


## What's Next?

This quick start covered the essentials. For more detailed guidance:

- **[Complete Workflow](workflow.html)**: Full analysis with interpretation 
  and visualization
- **[Methodology](methodology.html)**: Understanding the Bayesian Hybrid 
  Framework
- **[Dual Estimands](dual-estimands.html)**: Deep dive into estimands A, B, 
  and A*
- **[Diagnostics](diagnostics.html)**: Comprehensive convergence checking


## Summary Table

| Step | Function | Purpose |
|------|----------|---------|
| 1 | `compile_bhf_model()` | Compile Stan model (once per session) |
| 2 | `prepare_bhf_data()` | Transform data for Stan |
| 3 | `bhf_fit()` | Fit the Bayesian model |
| 4 | `variance_decomposition()` | Get ICC across estimands |
| 5 | `domain_estimates()` | Get domain-specific results |
| 6 | `overall_estimate()` | Get population-weighted overall |


---

*For questions or feedback, please visit the 
[GitHub repository](https://github.com/joonho112/bhfvar).*
